Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2018-02-16 paper-ieee-trans.tex

The practical implementation of the CI scenario leveraged on tightly integrated open
source tools such as GitHub \cite{github}, as the online source code repository,
Jenkins \cite{jenkins}, the event-response CI system, and Docker container
provisioning to provide instant computing power needed for executing the pipeline jobs.
The project defined a source code contribution workflow, based on GitHub Pull Requests
(PRs), where each change automatically triggers in Jenkins the associated quality and
metric tracking checks on PR creation or update. The set of required checks contained
in the CI pipeline, report their exit status back to GitHub preventing the change to be
approved if they are not successfully executed. As each change is validated, the chances
of early detection of defects increases. Within this scenario, the cost of defect solving
is dramatically reduced and the reliability of the software solutions improved, as any
bug or design issue is likely to be detected and subsequently corrected in this phase.

@@

\subsubsection{Continuous Delivery}
As DevOps suggests, frequent releases positively affect the reliability of the software as
they involve fewer changes and usually more correlated. The software updates of
{\sl INDIGO--DataCloud} products, taking place since the second major release, are passing through
a Continuous Delivery (CD) pipeline that adds the packaging of the software right after the
successful execution of the CI pipeline previously described. Consequently, CI-validated
software components are automatically delivered in online repositories without the need of
any human interaction but ultimately requiring expert supervision to validate the results.

{\sl INDIGO--DataCloud} software stack is delivered in form of Linux software packages (rpms, debs)
and Docker containers. The CD approach is different for each type of software packaging. \textit{Linux
package pipeline} intentionally delivers the packages produced in a pre-production -- preview --
branch. Before that, once the software packages have been created, they are uploaded to a
testing branch. The available automated deployment solution -- see \cite{subsec:sqa} -- checks
the installation and configuration of the component relying on the testing repository. On
successful completion, the packages are then automatically signed and subsequently moved to
the preview branch, from where they finally be made available in the production branch after
the release manager supervises the process.

@@

User application developers were provided with both a means to validate the
source code before merging and the creation of a new versioned Docker image,
automatically available in the {\sl INDIGO--DataCloud} application repository.

>>>>

The practical implementation of the CI scenario leveraged on tightly integrated open
source tools such as GitHub \cite{github}, as the online source code repository,

Jenkins \cite{jenkins}, the event-response CI system, and Docker container
provisioning to provide instant computing power needed for executing the pipeline jobs.
The project defined a source code contribution workflow, based on GitHub Pull Requests
(PRs), where each change automatically triggers in Jenkins the associated quality and
metric tracking checks on PR creation or update. The set of required checks contained
in the CI pipeline, report their exit status back to GitHub preventing the change to be
approved if they are not successfully executed. As each change is validated, the chances
of early detection of defects increases. Within this scenario, the cost of defect solving
is dramatically reduced and the reliability of the software solutions improved, as any
bug or design issue is likely to be detected and subsequently corrected in this phase.

@@

\subsubsection{Continuous Delivery}
As DevOps suggests, frequent releases positively affect the reliability of the software as
they involve fewer changes and usually are more correlated. The software updates of
{\sl INDIGO--DataCloud} products, taking place since the second major release, are passing through
a Continuous Delivery (CD) pipeline that adds the packaging of the software right after the
successful execution of the CI pipeline previously described. Consequently, CI-validated
software components are automatically delivered in online repositories without the need of
any human interaction but ultimately requiring expert supervision to validate the results.

{\sl INDIGO--DataCloud} software stack is delivered in form of Linux software packages (rpms, debs)
and Docker containers. The CD approach is different for each type of software packaging. \textit{Linux
package pipeline} intentionally delivers the packages produced in a pre-production -- preview --
branch. Before that, once the software packages have been created, they are uploaded to a
testing branch. The available automated deployment solution -- see Section \ref{subsec:sqa} -- checks
the installation and configuration of the component relying on the testing repository. On
successful completion, the packages are then automatically signed and subsequently moved to
the preview branch, from where they finally be made available in the production branch after
the release manager supervises the process.

@@

User application developers were provided with both a means to validate the
source code before merging and the creation of a new versioned Docker image,
automatically available in the {\sl INDIGO--DataCloud} applications repository.
